---
title: "GeneFishing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GeneFishing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GeneFishing)
data("pancreas")
data("insulin_secretion")

# log of count matrix
head(pancreas[, 1:4])

# GO term 
insulin_secretion[1:6]
```



## Probe for bait sets 
Search through ```insulin_secretion``` to find a subset of genes that can be used
as bait in GeneFishing.  Note, that in order for this vingette to compute quickly
```n_rounds = 10```, but in practice this should be at least 100.  

```{r probe_fishability}
set.seed(3908)
go_subset <- intersect(rownames(pancreas), insulin_secretion)
baits <- probeFishability(pancreas, go_subset, umap = FALSE, n_rounds = 10,
                          min_genes = 10)
baits
```

Will randomly sample genes and plot discovered bait with those randomly sampled
genes.  This shows two different random samples
```{r plot_probe_results1, fig.width=7, fig.height=4}
plot(baits)
```


```{r plot_probe_results2, fig.width=7, fig.height=4}
plot(baits)
```


## Do GeneFishing
Again, in practice one should do more than 25 rounds of GeneFishing, however to 
make the vignette compute in a reasonable amount of time we are only using 25
here.  Additionally, you can pass the output of the previous function, 
along with a index corresponding to the bait number you would like to do the 
fishing for.  

```{r gene_fishing}
fishing_results <- geneFishing(pancreas,
                               baits,
                               fishing_rounds = 25,
                               umap = FALSE)
fishing_results
```

Alternatively you can directly pass a vector of genes as bait.
```{r eval = FALSE}
# not run 
fishing_results <- geneFishing(pancreas,
                               baits$best_bait,
                               fishing_rounds = 25,
                               umap = FALSE)
```


Plot of capture frequency rate (CFR)
```{r plot_gene_fishing, fig.width=7, fig.height=4}
plot(fishing_results)
```

For a bait set with this tightness, one might consider using three clusters in GeneFishing
to get a more accurate set of fished out genes
```{r gf_three_clusters, fig.width=7, fig.height=4}
fishing_results_k3 <- geneFishing(pancreas,
                                  baits$best_bait,
                                  fishing_rounds = 25,
                                  k = 3,
                                  umap = FALSE)
fishing_results_k3
```


```{r plot_gf_three_clusters, fig.width=7, fig.height=4}
plot(fishing_results_k3)
```
