---
title: "GeneFishing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GeneFishing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## GeneFishing in simulated data
We simulate a dataset with two clouds of data.  The smaller cloud has 100 points, 
but we will say that we only know 20 of them belong to the group.  We will use
GeneFishing to recover the remaining 85 genes from the group. 


In the plot below we can see that in the first two features of ```X``` we have
no separation between the bait gene cloud and non-bait gene cloud.
```{r simulate_data, fig.height=4, fig.width=7}
library(GeneFishing)
library(MASS)
library(tidyverse)

set.seed(83)
# generate two random clouds of data
X <- rbind(mvrnorm(100, mu = sample(c(1:4), 10, replace = TRUE), 
                   Sigma = diag(rep(0.25, 10))), 
           mvrnorm(1000, mu = rep(1, 10), 
                   Sigma = diag(rep(1, 10))))

rownames(X) <- paste0("R", 1:nrow(X))

# use 20 of the smaller cloud as "bait"
bait <- sample(paste0("R", 1:100), 20)

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 color = ifelse(rownames(X) %in% bait, "bait", "non-bait"),
                 shape = ifelse(rownames(X) %in% paste0("R", 1:100), 
                                "bait cloud", "non-bait cloud")), 
             alpha = 0.5) + 
  theme_classic() + 
  theme(legend.title = element_blank()) + 
  scale_color_discrete(direction = -1) + scale_shape_discrete()
```

Instead of plotting the raw coordinates if we plot the spectral decomposition
using euclidean distance to compute similarity we can see the bait genes separating
from the non-bait.  GeneFishing leverages both the spectral decomposition and 
subsampling to cluster data.

```{r simulated_spectral_coordinates, fig.height=4, fig.width=7}
plotSpectralCoordinates(X, bait, umap = FALSE, method = "euclidean")
```

### Do GeneFishing
We can use ```GeneFishing``` to "fish-out" the 80 non-bait genes that are generated
in the bait gene cloud.  In the plot below we see the desired fishing output, which
show a small fraction of genes with capture frequency rate (CFR) close to 1 and 
most genes with CFR close to 0.  A CFR of 1 indicates that in every round of GeneFishing
that gene was in the cluster with the majority of the bait genes.

```{r simulated_gf, fig.height=4, fig.width=7}
fishing_results <- geneFishing(X, bait, method = "euclidean", alpha = 3,
                               fishing_rounds = 100, n_probing_rounds = 50)

plot(fishing_results)
```

We can also look at the fished out genes to see whether we recover the 80 genes
which were generated in the same cloud as the non-bait genes are recovered, and
how many genes in the other cloud are fished out (e.g. false-discoveries).  
```{r gf_simulated_fished_genes, fig.height=4, fig.width=7}
fished_genes <- fishing_results$fished_genes 

# genes in bait gene cloud that are fished out (but were not originally 
# labeled as bait)
bait_cloud <- setdiff(paste0("R", 1:100), bait)
intersect(fished_genes, bait_cloud) %>% length()

# genes not in non-bait gene cloud but are still fished out
intersect(fished_genes, paste0("R", 101:1100)) %>% length()
```

## GeneFishing in pancreas data
```{r setup}
data("pancreas")
data("insulin_secretion")

# log of count matrix
head(pancreas[, 1:4])

# GO term 
insulin_secretion[1:6] 
```



### Probe for bait sets 

We plot the spectral coordinates of all GO term genes in
the data below.  We can see that these genes are not separating from the non-bait
genes and thus GeneFishing with all GO term genes is not possible.

```{r GO_plot, fig.width=7, fig.height=4}
go_subset <- intersect(rownames(pancreas), insulin_secretion)
plotSpectralCoordinates(pancreas, go_subset, alpha = 3, 
                        umap = FALSE, method = "pearson")
```

Therefore we need to search through the  ```insulin_secretion``` GO term 
to find a subset of genes that can be used as bait in GeneFishing.  

Note, that in order for this vignette to compute quickly
```n_rounds = 10```, but in practice this should be at least 100.  

```{r probe_fishability}
baits <- probeFishability(pancreas, go_subset, umap = FALSE, n_rounds = 10,
                          min_genes = 10, method = "pearson")
baits
```

We plot the discovered bait with randomly sampled genes.  This shows two different 
random samples:
```{r plot_probe_results1, fig.width=7, fig.height=4}
plot(baits)
```


```{r plot_probe_results2, fig.width=7, fig.height=4}
plot(baits)
```

Notice that two of the discovered bait are not clustering with the other bait.  This
is not problematic, because the GeneFishing algorithm looks for genes which cluster
with the *majority* of the bait. 

### Do GeneFishing
Again, in practice one should do more than 25 rounds of GeneFishing, however to 
make the vignette compute in a reasonable amount of time we are only using 25
here.  You can pass the output of the previous function, 
along with a index corresponding to the bait number you would like to do the 
fishing for.  In this case the provided similarity ```method``` will be ignored
and the method from ```probeFishability``` will be used.

```{r gene_fishing}
fishing_results <- geneFishing(pancreas,
                               baits,
                               fishing_rounds = 25,
                               umap = FALSE)
fishing_results
```

Alternatively you can directly pass a vector of genes as bait, in this case you
will need to provide the similarity ```method``` if you don't want to use the default 
of Spearman's correlation.  

```{r eval = FALSE}
# not run 
fishing_results <- geneFishing(pancreas,
                               baits$best_bait,
                               fishing_rounds = 25,
                               umap = FALSE,
                               method = "pearson")
```


Plot of capture frequency rate (CFR)
```{r plot_gene_fishing, fig.width=7, fig.height=4}
plot(fishing_results)
```

For a bait set with this tightness, one might consider using three clusters in GeneFishing
to get a more accurate set of fished out genes
```{r gf_three_clusters, fig.width=7, fig.height=4}
fishing_results_k3 <- geneFishing(pancreas,
                                  baits$best_bait,
                                  fishing_rounds = 25,
                                  k = 3,
                                  umap = FALSE,
                                  method = "pearson")
fishing_results_k3
```


```{r plot_gf_three_clusters, fig.width=7, fig.height=4}
plot(fishing_results_k3)
```
